API design for machine learning software: experiences from the scikit-learn project

### 2013-06-26 paper.tex
spell check using aspell

In other places of the library, in particular in the vectorization code
for unstructured input, the toolkit is also designed to be
extensible. Here, estimators provide hooks for user-defined code: objects or
functions that provide a specific API can be given as arguments at vectorizer
construction time. The library then calls into this code, communicating with it by passing objects of standard Python/Numpy types.
Again, such external user code can be kept agnostic of the \sklearn
class hierarchy.

@@

Our implementation guidelines emphasize writing efficient but readable code.
In particular, we focus on making the codebase easily maintainable and
understandable in order to favour external contributions. Whenever practicable,
algorithms implemented in \sklearn are written in Python,
using Numpy vector operations for numerical work.
This allows for the code to remain concise, readable and
efficient. For critical algorithms that cannot be easily and efficiently
expressed as NumPy operations, we rely on Cython \citep{behnel2011cython}
to achieve decent performance and scalability. Cython is a
compiled programming language that extends Python with static typing. It
produces efficient C extension modules that are directly importable from the
Python run-time system. Examples of algorithms written in Cython include
stochastic gradient descent for linear models, some graph-based clustering
algorithms and decision trees.

To facilitate the installation and thus adoption of \sklearn,
the set of external dependencies is kept to a bare minimum:
only Python, NumPy and SciPy are required for a functioning installation.
Binary distributions of these are available for the major platforms.
Visualisation functionality depends on Matplotlib \citep{hunter2007matplotlib}
and/or Graphviz \citep{gansner2000},
but neither is required to do machine learning.
When feasible, external libraries are integrated into the codebase.
In particular, \sklearn includes modified versions of \textsf{LIBSVM} and \textsf{LIBLINEAR}
\citep{chang2011libsvm,fan2008}, both written in C{}\verb!++!
and wrapped using Cython modules.

@@

At a lower-level, parallel processing is a potential point of improvement.
Some estimators in \sklearn are already able to leverage multi-core processors,
but only in a coarse-grained fashion.
At present, parallel processing is difficult to accomplish in the Python environment;
\sklearn targets the main implementation, CPython,
which cannot execute Python code on multiple CPUs simultaneously.
It follows that any parallel task decomposition must either be done
inside Cython modules,
or at a level high enough to warrant the overhead
of creating multiple OS-level processes,
and the ensuing inter-process communication.
Parallel grid search is an example of the latter approach
which has already been implemented.
Recent versions of Cython include support for the OpenMP standard
\citep{dagum1998openmp},
which is a viable candidate technology
for more fine-grained multicore support in \sklearn.

@@

\subsection*{Acknowledgements}

The authors and contributors acknowledge active support from INRIA. Past and
present sponsors of the project also include Google for funding
scholarships through its Summer of Code program,
the Python Software Fundation and Tinyclues for funding coding sprints.

>>>>

In other places of the library, in particular in the vectorization code
for unstructured input, the toolkit is also designed to be
extensible. Here, estimators provide hooks for user-defined code: objects or
functions that provide a specific API can be given as arguments at vectorizer
construction time. The library then calls into this code, communicating with it by passing objects of standard Python/NumPy types.
Again, such external user code can be kept agnostic of the \sklearn
class hierarchy.

@@

Our implementation guidelines emphasize writing efficient but readable code.
In particular, we focus on making the codebase easily maintainable and
understandable in order to favor external contributions. Whenever practicable,
algorithms implemented in \sklearn are written in Python,
using NumPy vector operations for numerical work.
This allows for the code to remain concise, readable and
efficient. For critical algorithms that cannot be easily and efficiently
expressed as NumPy operations, we rely on Cython \citep{behnel2011cython}
to achieve decent performance and scalability. Cython is a
compiled programming language that extends Python with static typing. It
produces efficient C extension modules that are directly importable from the
Python run-time system. Examples of algorithms written in Cython include
stochastic gradient descent for linear models, some graph-based clustering
algorithms and decision trees.

To facilitate the installation and thus adoption of \sklearn,
the set of external dependencies is kept to a bare minimum:
only Python, NumPy and SciPy are required for a functioning installation.
Binary distributions of these are available for the major platforms.
Visualization functionality depends on Matplotlib \citep{hunter2007matplotlib}
and/or Graphviz \citep{gansner2000},
but neither is required to do machine learning.
When feasible, external libraries are integrated into the codebase.
In particular, \sklearn includes modified versions of \textsf{LIBSVM} and \textsf{LIBLINEAR}
\citep{chang2011libsvm,fan2008}, both written in C{}\verb!++!
and wrapped using Cython modules.

@@

At a lower-level, parallel processing is a potential point of improvement.
Some estimators in \sklearn are already able to leverage multicore processors,
but only in a coarse-grained fashion.
At present, parallel processing is difficult to accomplish in the Python environment;
\sklearn targets the main implementation, CPython,
which cannot execute Python code on multiple CPUs simultaneously.
It follows that any parallel task decomposition must either be done
inside Cython modules,
or at a level high enough to warrant the overhead
of creating multiple OS-level processes,
and the ensuing inter-process communication.
Parallel grid search is an example of the latter approach
which has already been implemented.
Recent versions of Cython include support for the OpenMP standard
\citep{dagum1998openmp},
which is a viable candidate technology
for more fine-grained multicore support in \sklearn.

@@

\subsection*{Acknowledgments}

The authors and contributors acknowledge active support from INRIA. Past and
present sponsors of the project also include Google for funding
scholarships through its Summer of Code program,
the Python Software Foundation and Tinyclues for funding coding sprints.
